Late Pleistocene stone artefact technology at Mau A, Yen Bai, Norther Vietnam

### 2022-02-17 analysis/paper/paper.Rmd
Added PCA code to create dorsal cortex biplots

<!--QH: Just putting all of the code and writing here. Not going to worry about formatting for now -->
<!--QH: On a secondary note, I managed to cause an error with docker in record time. We will need to fix that later-->
<!--QH: Lastly, I'm going to do my best to get the code working tonight. If that fails, I will still write up my interpretation using prior images of the visualization-->

@@

<!--QH: I have been fiddling about with code and github for awhile. I plan on fixing everything when I wake up, and, failing that, Friday afternoon.-->

```{r}
library(Ckmeans.1d.dp)

test_result_dorsal_cortex <- Ckmeans.1d.dp(ma_flakes_tr$dorsal_cortex_perc)

# plot(test_result_dorsal_cortex)

# It seems that there is no clear clustering in the Dorsal Cortex values. I will have to use values from the literature to divide the cortex values

# try mass:
test_result_mass <- Ckmeans.1d.dp(ma_flakes_tr$mass)

# plot(test_result_mass)
# we got six clusters
```

```{r}
ma_flakes_tr_with_groups <-
  ma_flakes_tr %>%
  mutate(dorsal_cortext_group = as.factor(
           case_when(
             between(dorsal_cortex_perc, 0, 0) ~ "Tertiary",
             between(dorsal_cortex_perc, 1, 50) ~ "Secondary",
             between(dorsal_cortex_perc, 51, 100) ~ "Primary"
           ))) %>% 
 mutate(mass_group = as.factor(test_result_mass$cluster))

# plot mass groups
ggplot(ma_flakes_tr_with_groups)  +
  aes(mass,
      fill = mass_group) +
  geom_histogram() +
  scale_fill_viridis_d(name = "Mass\ncluster") +
  xlab("Mass (g)") +
  theme_minimal(base_size = 14) +
  theme(panel.background = element_rect(fill='white',
                                        colour = "white"),
        plot.background = element_rect(fill='white',
                                       colour = "white"))

>>>>

## PCA Analysis on Dorsal Cortex Groups

@@

```
```{r}

library(tidyverse)
library(geomorph)
library(Morpho)
library(mclust)
library(cluster)
library(factoextra)

lands <- readland.tps(
  here::here("data/derived_data/landmarks.tps"),
  specID = "ID")

# drop outliers  identified in the previous script
outlier_ids <- c(1152,  606, 1147, 1125,  756 ,1023,  457,
                 813,  415,  877,  677, 1055,  239,  811,  601,  861,
                  794,  103 , 598 ,1110 , 906 ,  59 ,   4)

lands_no_outliers <- lands[, , -outlier_ids]
lands_no_outliers_names <- dimnames(lands_no_outliers)[[3]]

sup.geo <- gpagen(lands_no_outliers,  ProcD=F, Proj = T)

# Principal component analysis (PCA) using a base R function
pca.geo <- prcomp(two.d.array(sup.geo$coord))

# TODO: find a meaningful way to break this up into groups
# pass these groups onto the PCA for drawing ellipses

# this block of code below shows one way to discretize cortex percentage
# into three groups. We need to explore other cutoffs, and other grouping
# variables besides dorsal cortex. This code below is a good template for
# starting these explorations
ma_flakes_tr_with_groups <-
  ma_flakes_tr %>%
  slice(-outlier_ids) %>%
  mutate(dorsal_cortext_group =
           case_when(
             between(dorsal_cortex_perc, 0, 0) ~ "Tertiary",
             between(dorsal_cortex_perc, 11, 60) ~ "Secondary",
             between(dorsal_cortex_perc, 61, 100) ~ "Primary"
           ))

# ggplot2-based plot of the PCA results using a factoextra function
fviz_pca_ind(pca.geo,
             geom.ind = "point", # show points without labels
            #  line below sets the grouping column for colouring the points
             habillage = ma_flakes_tr_with_groups$dorsal_cortext_group,
             palette = "npg", #color palette, adjust to preferences
             addEllipses = T, #adds ellipses or convex hulls
             ellipse.type = "convex", #convex hulls
             mean.point = F, #don’t show average point
             legend.title = "SP") +  #adjust accordingly)
             coord_fixed(ratio = 1) #adjustable for variation-weighted plots
